### 2023-09-15 paper-correlations.tex
Improvements from Alexey

\documentclass[1p]{elsarticle}
% Latex
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
% Xelatex
%\usepackage{polyglossia}
%\usepackage{fontspec}
%\setdefaultlanguage{english}
%\setmainfont{DejaVu Sans}
%\setsansfont{DejaVu Sans}
%\setmonofont{DejaVu Sans Mono}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{bm}
\usepackage{hyperref}
\usepackage{cleveref}
\usepackage{xcolor}
\usepackage{algpseudocode}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage[export]{adjustbox}

\newtheorem{definition}{Definition}

\definecolor{light-gray}{gray}{0.95}
\newcommand{\code}[1]{\colorbox{light-gray}{\texttt{#1}}}

\bibliographystyle{elsarticle-num}
\journal{Computational Material Science}

\begin{document}

\author[Smith]{Vasily~Postnicov}
\author[Smith,MIT]{Aleksei~Samarin}
\author[Smith]{Marina ~V. Karsanina}
\author[Aus]{Mathieu ~Gravey}
\author[MIPT]{Aleksey~Khlyupin}
\author[Smith]{Kirill ~M. Gerke}
\ead{kg@ifz.ru}

\address[Smith]{Schmidt Institute of Physics of the Earth of Russian Academy of
  Sciences, Moscow, 107031, Russia}
\address[MIT]{Computational Mathematics and Cybernetics, Lomonosov Moscow State
  University, Moscow, 119991, Russia}
\address[Aus]{Institute for Interdisciplinary Mountain Research, Austrian
  Academy of Sciences, Innsbruck, 6020, Austria}
\address[MIPT]{Moscow Institute of Physics and Technology, Institutskiy Pereulok 9,
Dolgoprudny, Moscow, Russia 141700}

\title{Evaluation of classical correlation functions from 2/3D images on CPU and
  GPU architectures: introducing CorrelationFunctions.jl}

\begin{abstract}
Correlation functions are becoming one of the major tools for quantification of
structural information that is usually represented as 2D or 3D image of porous
material.  In this paper we introduce \code{CorrelationFunctions.jl} open-source
package developed in Julia and capable of computing all classical correlation
functions based on imaging input data. Images include both binary and
multi-phase representations.  Our code is capable of evaluating two-point
probability $S_2$, phase cross-correlation $\rho_{ij}$, cluster $C_2$,
lineal-path $L_2$, surface-surface $F_{ss}$, surface-void $F_{sv}$, pore-size
$P$ and chord-length $p$ distribution functions on both CPU and GPU
architectures.  Where possible, we presented two types of computations: full
correlation map (correlations of each point with other points on the image, that
also allows obtaining ensemble averaged CF) and directional correlation
functions (currently in major orthogonal and diagonal directions).  Such an
implementation allowed for the first time to assemble a completely free solution
to evaluate correlation functions under any operating system with well
documented application programming interface (API). Our package includes
automatic tests against analytical solutions that are described in the paper. We
measured execution times for all CPU and GPU implementations and as a rule of
thumb full correlation maps on GPU are faster than other methods. However, full
maps require more RAM and, thus, are limited to available RAM resources. On the
other hand, directional CFs are memory efficient and can be evaluated for huge
datasets -- this way they are the first candidates for structural data
compression of feature extraction. The package itself is available through Julia
package ecosystem and on GitHub
(\url{https://github.com/fatimp/CorrelationFunctions.jl}), the latter source
also contains documentation and additional helpful resources such as tutorials.
We believe that a single powerful computational tool such as
\code{CorrelationFunctions.jl} presented in this paper will significantly
facilitate the usage of correlation functions in numerous areas of structural
description and research of porous materials, as well as in machine learning
applications. We also present some example as applied to ceramic, soil composite
and oil-bearing rock samples based on their 3D X-ray tomography and 2D scanning
electron microscope images. Finally, we conclude our paper with discussion of
possible ways to further improve presented computational framework.
\end{abstract}

\maketitle

\section{Background and motivation}
\label{sec:background}
Heterogeneous materials are ubiquitous both in nature and in industrial human
activities, and their physical properties are interconnected to structure
\cite{Torquato_book,Sahimi_book}. The information about the structure usually
comes in the way of 2D and 3D images obtained by microscopy
\cite{moussaoui2018,neumann2019,FIB-SEMpaper}, tomography \cite{xctmat_book} and
other related methodologies.  This structural information can be used for
numerous types of analysis, including simulations of different processes within
studied materials or evaluation of their properties based on 3D structural data
\cite{youssef2005,Miao2017,FDMSS}. Unlike computational modelling techniques a
whole class of approximate methods exists including (rigorous) bounds that
allows very fast estimations of physical properties
\cite{eshelby1957,berryman1986use,rozanski2023} albeit the bounds themselves are
too wide to be applicable in the majority of practical problems. Nonetheless, it
is noteworthy that many of such bounds are originally based on structural
descriptors in the form of correlation functions (CFs).

Two ways exist to obtain correlation functions for a given structure at hand:
measure them either experimentally with the help of, for example, scattering
intensity \cite{debye1957scattering,li2018direct} or from digital images
\cite{berryman1985measurement,ma2018SS}. None is perfect, as the first approach
suffers from the limitation of CFs that can be obtained this way, while the
second one provides information with limited resolution or/and resolution to
field-of-view ratio \cite{gerke2015universal}. In addition to the resolution
conundrum, the most useful imaging methods such as X-ray computed tomography
(XCT) and scanning electron microscopy (SEM) provide gray-scale images (e.g.,
X-ray attenuation or electron back-scattering distributions) due to their
underlying physical principles and require labeling of constituent phases (or
segmentation) before computation of CFs \cite{samarin2023robust}. If properly
segmented \cite{NNseg}, high-resolution digital images do provide a possibility
to compute any correlation function and, thus, possibility to address numerous
fundamental and practical research problems.

Correlation functions as invaluable universal descriptors of structure are
utilized in a multitude of scientific disciplines: material sciences
\cite{hasanabadi20163d,Havelka,feng2018reconstruction,chen2022quantifying}, rock
physics \cite{ledesma2018effect}, soil physics and hydrology
\cite{PLoS_ONE,KarsaninaEJSS}, cosmology and food engineering
\cite{TakadaJain,Derossi2019}, biology \cite{veatch2012correlation} and many
others. Computed from 2D and 3D images, CFs can be applied to:
\begin{enumerate}
\item characterize the morphology and representativeness via correlation
  lengths \cite{vcapek2009stochastic,thovert2011grain,tensorPRE};
\item perform stochastic reconstructions from experimentally measured CFs
  or 2D to 3D reconstructions based on CFs that can be extracted from less
  dimensions or through penetrable sphere model
  \cite{Adler_recon,Y-T,tahmasebiPRL,hasanabadi20163d,karsaninaPRL};
\item compare different structures to each other, including verification of
  stochastic reconstructions \cite{vcapek2009stochastic,tahmasebiPRL,EPL2};
\item compress structural information in the form of raw or parameterized CFs
  with the possibility to recover the structure using stochastic
  reconstruction \cite{SciRep1,Havelka,KarsaninaEJSS};
\item describe structural dynamics under different boundary conditions
  \cite{chen2015dynamic,xu2022correlation,fomin2023soil};
\item extract structural features for deep learning
  \cite{pilania2017multi,kamrava2020linking,roding2020predicting,KarsaninaEJSS};
\item fuse multi-scale images and structural information as obtained by
  different methods into a single digital model \cite{Geoderma2018};
\item quantify spatial heterogeneity \cite{zhang2000pore,REVpaper}.
\end{enumerate}

Stochastic reconstruction is a separate huge topic, as this approach allows to
solve an inverse problem and recover structure from known set of correlation
functions
\cite{Adler_recon,Y-T,vcapek2009stochastic,hasanabadi20163d,Havelka,feng2018reconstruction,tahmasebiPRL,EPL2,thovert2011grain,tensorPRE}. This
ability for recovery serves as the basis for majority of usages in the list
above. While it is necessary to compute CFs from digital images as a target set
for stochastic reconstructions, they rely more on efficient recomputations or
optimizations, e.g., re-evaluation during simulated annealing. Such
optimizations are not the part of the computational package described here and
are not considered in this work.

There are different types of correlation functions that in general describe some
probabilities. The main parameter of any correlation function is the number of
points that are utilized to evaluate such a probability. While so-called
$n$-point probability function \cite{Torquato_book} will completely describe any
structure in the limit of $n \rightarrow$ number of voxels/pixels on the image,
computing or storing such a function is not practical. Based on information
content, it was shown that increasing the number of point $n > 2$ only
marginally improves the quality of the structure quantification and these
improvements decay with increasing $n$ \cite{yao1993high,Gommes2}. For
stochastic reconstruction purposes computation of higher-order statistics will,
arguably, not be balanced by increased number of points, but this is the topic
of active research \cite{skolnick2021,cherkasovIC}. For aforementioned reasons,
we mainly focus on 2-point statistics, leaving the possibility to include higher
order CFs in the future work. The simplest 2-point probability function ($S_2$
or autocorrelation) actually arise from small angle scattering experiments and
measures the probability that both ends of line segment with a given length fall
into the same phase. Other types include the probability of a whole line segment
to fall into the phase ($L_2$ function) or ``pointing'' with its ends to the
same cluster ($C_2$ function) and mainly originate from aforementioned bounds on
physical properties such as permeability and elasticity. The general idea of
computation of different 2-point CFs is shown in \cref{fig:functions}. All major
details and numerous analytical cases can be found in seminal work of Torquato
\cite{Torquato_book}. To all CFs described in this book and applicable to
digital images we shall refer to as classical correlation functions.

\begin{figure}[ht]
  \centering
  \includegraphics[width=0.9\linewidth]{images/cf_1.png}
  \caption[]{A schematic representation of a binary porous media (pores are
    shaded) with examples of CFs computations (all classical correlation
    functions are presented according to the legend on the right
    side). Computations take into account all possible configurations of a line
    segment (or a N-dimensional ball for pore size function) which either make a
    contribution to the CF (conventionally shown as ``CF = 1'') or do not make a
    contribution (``CF = 0'').}
  \label{fig:functions}
\end{figure}

While evaluation of correlation functions from 2D and 3D images was a topic of
numerous research papers, the information is highly fragmented, especially in
terms of algorithms and their implementation in the code. Some snippets of code
are available for some functions, but they lack general API and are usually
implemented in proprietary interpreted languages (e.g. Matlab). Due to a recent
leap forward in GPU-based computations, some papers describe computations on
this architecture, but the code is usually not provided. All in all, the
efficient and open-source single API solution utilizing both CPU and GPU
architectures is currently absent.

The aim of this paper is to establish a general open-source solution allowing to
compute all classical correlation functions. This solution should be well
documented with all computational algorithms explained, it should also leverage
on recent advances in programming high-intensity computations and utilize both
CPU and GPU power. Our answer in the form of \code{CorrelationFunctions.jl}
package allows to compute numerous CFs easily on any operating system. The
manuscript describing the package is organized as follows: in \cref{sec:math} we
introduce a reader to the most common and useful correlation functions,
\cref{sec:map} presents all algorithmic details for computing full CFs maps (map
method, full 2-point statistics for a given image). In \cref{sec:scan} we
discuss the second method to calculate correlation functions which is called the
scan approach. In \cref{sec:verification} we verify our algorithms based on
known analytical solutions. In \cref{sec:examples} some application examples are
provided, including CFs evaluation for multi-phase materials. Section
\ref{sec:efficiency} deals with computational efficiency of our code, we compare
computational times for 2D and 3D images of different sizes for scanning and map
methods as evaluated using either CPU or GPU.  Finally, sections
\ref{sec:outlook} and \ref{sec:summary} summarize all results and outline
possible future improvements.

\section{A brief primer on correlation functions}
\label{sec:math}
In this section we shall provide major properties and definitions of correlation
functions supported by our package; the full list of functions is (see
\cref{fig:functions} for graphical explanation in addition to definitions
below):
\begin{itemize}
\item Two-point probability function $S_2$,
\item Lineal-path function $L_2$,
\item Cluster function $C_2$,
\item Surface-surface function $F_{ss}$,
\item Surface-void function $F_{sv}$,
\item Pore size function $P$,
\item Chord length function $p$,
\item Phase cross-correlation function $\rho_{ij}$.
\end{itemize}
Note that cross-correlation $\rho_{ij}$ is actually not a classical CF, but
technically is a variety of $S_2$ for the case when the end of line segment lie
in different phases. This function is useful for multi-phase structures only, as
for binary media two-point probability CFs are interdependent.

The most widely used is a two-point probability correlation function $S_2$, or
autocorrelation. A definition for homogenous media is as follows:
\begin{definition}
  Two point function $S_2^{(i)}(\bm{r})$ equals to probability that both ends
  of a vector $\bm{r}$ lie in phase $i$ when the vector is randomly thrown
  into the sample.
\end{definition}
For isotropic media a vector $\bm{r}$ can be reduced to a scalar value
$r$ (i.e., a correlation function depends only on length of a vector and not on
its direction). Mathematically this definition can be expressed with the
following equation:
\begin{equation}
  S_2^{(i)}(\bm{r}) = \langle I^{(i)}(\bm{x}) I^{(i)}(\bm{x} +
  \bm{r}) \rangle
  \label{eq:s2-def}
\end{equation}
where $I^{(i)}$ is an indicator function for a set of all points belonging to
the phase $i$ and $\langle \cdots \rangle$ is ensemble average.

A function closely related to $S_2$ is phase cross-correlation function
$\rho_{ij}$ which is defined as follows:
\begin{equation}
  \rho_{ij}(\bm{r}) = \langle I^{(i)}(\bm{x}) I^{(j)}(\bm{x} +
  \bm{r}) \rangle
  \label{eq:cross-def}
\end{equation}

Another basic correlation function is a lineal path function $L_2$. It provides
some non-trivial information (albeit usually not complete
\cite{vcapek2011transport}) about the connectedness of the phase and is defined
for homogenous media as:
\begin{definition}
  Lineal path function $L_2^{(i)}(\bm{r})$ equals to probability that a
  vector $\bm{r}$ lies wholly in phase $i$ when randomly thrown into the
  sample.
\end{definition}
If a sample is isotropic a vector $\bm{r}$ can be replaced with its length
$r$ and the definition becomes:
\begin{definition}
  Lineal path function (for isotropic media) $L_2^{(i)}(r)$ equals to
  probability that a line segment of length $r$ lies wholly in phase $i$ when
  randomly thrown into the sample.
\end{definition}

The function that contain a lot of structural connectivity information \cite{JiaoPNAS}
is cluster function $C_2$.
\begin{definition}
  Cluster function $C_2^{(i)}(\bm{r})$ equals to probability that both
  ends of a vector $\bm{r}$ lie in the same cluster when the vector is
  randomly thrown into the sample. A cluster is a set of points of the same
  phase $i$ where any two points can be connected by a path lying entirely in
  that phase.
\end{definition}
Note that for a fully connected phase $C_2$ will be equal to $S_2$, something
we shall utilize later for computations.

The next two functions are called surface-surface ($F_{ss}$) and surface-void
($F_{sv}$) correlation functions and describe an interphase between phases. We
skip strict verbal definitions and first introduce an interface indicator
function instead:
\begin{equation*}
  M^{(i)}(\bm{r}) = | \nabla I^{(i)}(\bm{r}) |
\end{equation*}
Here differentiation is understood in the sense of generalized functions,
because an ordinary differential of $I$ is either zero or undefined. For
homogenous media the definitions of surface functions then are:
\begin{align}
  F_{ss}^{(i)}(\bm{r}) &= \langle M^{(i)}(\bm{x}) M^{(i)}(\bm{x} +
  \bm{r}) \rangle \label{eq:fss-def} \\
  F_{sv}^{(i)}(\bm{r}) &= \langle M^{(i)}(\bm{x}) I^{(void)}(\bm{x}
  + \bm{r}) \rangle \label{eq:fsv-def}
\end{align}
These functions are defined for media with dimensionality of 2 and
higher. $F_{sv}$ is a physical quantity inversely proportional to length and
$F_{ss}$ is inversionaly proportional to surface area. As always, there are
versions of these functions applicable to scalar argument for isotropic media.

Pore size function $P(r)$ is a probability density function defined as follows:
\begin{definition}
  $P(r)dr$ equals to the probability that a randomly chosen point in a set of
  points belonging to the void phase lies at a distance between $r$ and $r + dr$
  from the nearest point on the pore-solid interface.
\end{definition}

Finally, a chord length function $p^{(i)}(r)$ is defined as:
\begin{definition}
  $p^{(i)}(r)dr$ equals to the probability of finding a chord of length between
  $r$ and $r+dr$ in phase $i$. Chord is a line segment which lies entirely in
  the same phase and touches the interface with its ends.
\end{definition}

More detailed information on all classical CFs can be found in
\cite{Torquato_book}.

\section{Computational methodology}
\code{CorrelationFunctions.jl} is capable to compute all aforementioned
correlation functions. Where possible we provide two implementations for each
CF. The first implementation computes a correlation function in specific
predefined directions (i.e., orientation of test vectors thrown into a sample is
fixed) \cite{jiao2014chawla,EPL1}.  There are one predefined direction in 1D
case, four directions in 2D case (two orthogonal and two diagonal) and thirteen
directions in 3D case (three orthogonal, six diagonal within each orthogonal
plane and four diagonal). All directions available are schematically represented
in \cref{fig:directions}. The line segment scanning approach is slower for some
CFs but requires much less memory resources. The full map approach establishes
correlations between all points on the image and is faster. Both implementations
are contained in \code{Directional} and \code{Map} modules of the package,
correspondingly.

With the help of the package one can compute CFs in periodic and non-periodic
modes.  Periodic mode assumes that an input sample is periodically continued to
the whole Euclidean space. Non-periodic mode treats the input as zero-padded image.

\begin{figure}[ht]
  \centering
  \includegraphics[width=0.9\linewidth]{images/cf_diag.png}
  \caption[]{Available directions to compute directional correlation functions
    in 2D and 3D cases with line scanning approach.}
  \label{fig:directions}
\end{figure}

\subsection{Full correlation map}
\label{sec:map}
Definition \cref{eq:s2-def} can be rewritten as follows:
\begin{equation}
  S_2^{(i)}(r) = \frac{F^{-1} [|F [A^{(i)}](z)|^2] (r)}{N(r)} \label{eq:s2-ft}
\end{equation}
where $A^{(i)} \xleftarrow{I^{(i)}} A$ is a binary array obtained by element-wise
application of indicator function $I^{(i)}$ to an input $A$, $F$ is a discrete
Fourier transform (DFT) operator and $N(r)$ is a total number of trials (vectors
thrown into the sample). If we want to calculate $S_2$ function in periodic mode
we need only to apply \cref{eq:s2-ft} to an input to obtain the result. In
non-periodic mode we need to pad an input with zeros to at least twice the size
minus one for each dimension. In periodic mode a total number of trials is
independent of $r$ and equals to the number of elements in an input array. In
non-periodic mode $N(r)$ is calculated as described in
\cref{sec:number-of-trials}. The following algorithm summarizes all the above:
\begin{algorithmic}[1]
  \Procedure{$S_2$}{$A, phase, periodic$}
  \If{$\neg periodic$}
  \State $A \gets zeropad(A)$
  \Comment Pad with at least $2\cdot size(A, i) - 1$ zeros for each dimension $i$.
  \EndIf
  \State $A^{(phase)} \gets I^{(phase)} (A)$
  \State $\hat{A}^{(phase)} \gets F[A^{(phase)}]$
  \State $\hat{S}_2 \gets |\hat{A}^{(phase)}|^2 / N$
  \Comment $N$ is an array with numbers of trials
  \State \textbf{return} $F^{-1} [\hat{S}_2]$
  \EndProcedure
\end{algorithmic}
Algorithmic complexity of this implementation is $O(M \log M)$, where $M$ is the
number of elements in the input image.

Computation of $S_2$ is a special case of a cross-correlation function $\rho_{ij}$
which computes a correlation between phases $i$ and $j$ using a formula:
\begin{equation}
  \rho_{ij}(r) = \frac{F^{-1} [F[A^{(i)}](z) \overline{F[A^{(j)}](z)}] (r)}{N(r)} \label{eq:cross-ft}
\end{equation}
that algorithmically operates as follows:
\begin{algorithmic}[1]
  \Procedure{$\rho_{ij}$}{$A, i, j, periodic$}
  \If{$\neg periodic$}
    \State $A \gets zeropad(A)$
  \EndIf
  \State $A^{(i)} \gets I^{(i)} (A)$
  \State $A^{(j)} \gets I^{(j)} (A)$
  \State $\hat{A}^{(i)} \gets F[A^{(i)}]$
  \State $\hat{A}^{(j)} \gets F[A^{(j)}]$
  \State $\hat{\rho} \gets \hat{A}^{(i)} \overline{\hat{A}^{(j)}} / N$
  \State \textbf{return} $F^{-1} [\hat{\rho}]$
  \EndProcedure
\end{algorithmic}

Cluster function can be calculated with the help of previuously described
procedures.  Firstly, we apply an indicator function to an input $A$:
$A^{(i)} \xleftarrow{I^{(i)}} A$. Then we label clusters in $A^{(i)}$ using a
function that is a part of Julia \code{ImageMorphology.jl} library.  Labeling
algorithm must take into account if the function is calculated with periodic
boundary conditions \cite{evstigneev2023}. For periodic boundary conditions we
implemented our own solution as part of the package, as it is absent in the
public domain. Then $C_2$ function is a sum of $S_2$ functions calculated for
all clusters separately. Algorithmic complexity of this implementation is
$O(MC\log M)$ where $M$ is the number of elements in the input and $C$ is the
number of clusters. The algorithm is then written as:
\begin{algorithmic}[1]
  \Procedure{$C_2$}{$A, phase, periodic$}
  \State $A^{(phase)} \gets I^{(phase)} (A)$
  \State $CC \gets label\_components(A^{(phase)}, periodic)$
  \State $N \gets maximum(CC)$ \Comment Number of connected components
  \State \textbf{return} $\sum\limits_{n=1}^N S_2(CC, n, periodic)$
  \EndProcedure
\end{algorithmic}

For surface-surface and surface-void functions we use an edge detection filter
\cite{samarin2023robust} and then calculate either autocorrelation of the edge
or cross-correlation of the edge and the void phase. The edge is extracted by
convolving an input with a short high-pass filter $H$. It has a width of 7
pixels/voxels and all its coefficients with exception of the central coefficient
are inversely proportional to distance from the center:
\begin{equation}
  \begin{aligned}
    H_k &= S \left\{
    \begin{array}{ll}
      -\sum\limits_{\substack{l \\ l \ne 0}} H_l & \quad k = 0 \\
      1 / \rho(k, 0) & \quad \text{otherwise}
    \end{array}
    \right. \\
    k &\in \{-3, -2, \dots, 3\}^D
  \end{aligned}
  \label{eq:filter-7x7}
\end{equation}
where $S$ equals to $30.45849$ in 2D and $172.96232$ in 3D case. The parameters
of the filter were adjusted based on comparison against exact computations of
surface functions on smooth-boundary sets \cite{postcov23}.

An algorithm for $F_{ss}$ function is the following:
\begin{algorithmic}[1]
  \Procedure{$F_{ss}$}{$A, phase, periodic$}
  \If{$\neg periodic$}
    \State $A \gets zeropad(A)$
  \EndIf
  \State $A^{(phase)} \gets I^{(phase)} (A)$
  \State $A_H \gets H * A^{(phase)}$
  \State $\hat{A}_H \gets F[A_H]$
  \State $\hat{F}_{ss} \gets |\hat{A}_H|^2 / N$
  \State \textbf{return} $F^{-1} [\hat{F}_{ss}]$
  \EndProcedure
\end{algorithmic}

Using \cref{eq:cross-ft} we obtain an algorithm for surface-void function:
\begin{algorithmic}[1]
  \Procedure{$F_{sv}$}{$A, phase, periodic$}
  \If{$\neg periodic$}
    \State $A \gets zeropad(A)$
  \EndIf
  \State $A^{(phase)} \gets I^{(phase)} (A)$
  \State $A^{(void)} \gets I^{(void)} (A)$
  \State $A_H \gets H * A^{(phase)}$
  \State $\hat{A}_H \gets F[A_H]$
  \State $\hat{A}^{(void)} \gets F[A^{(void)}]$
  \State $\hat{F}_{sv} \gets \hat{A}_H \overline{\hat{A}^{(void)}} / N$
  \State \textbf{return} $F^{-1} [\hat{F}_{sv}]$
  \EndProcedure
\end{algorithmic}

The interface between phases must not change rapidly for the edge filter to
work correctly. Therefore, the algorithms for surface functions require input
images to have apropriate resolution. A criterion for goodness can be expressed with
$C_{\alpha}$ parameter. Let $\alpha \in [0, 1]$ and $f(x)$ be
(band-limited) input image ($x \in \mathbb{R}^2$ or $\mathbb{R}^3$). We define
$C_\alpha$ as follows:
\begin{align*}
  f_0(x) &= f(x) - \langle f(x) \rangle \\
  C_\alpha &= \frac{\int_{-\alpha\omega}^{\alpha\omega} |\hat{f_0}(z)|^2
    dz}{\int_{-\omega}^{\omega} |\hat{f_0}(z)|^2 dz}
\end{align*}
where $\hat{f_0}$ is a Fourier transform of $f_0$ and $\omega$ is a folding
frequency which depends only on resolution of the image. We have previously
shown that criterion $C_{0.5} > 0.97$ holds then the image is suitable for
robust $F_{ss}$ and $F_{sv}$ evaluation \cite{samarin2023robust}.  Note that
function \code{lowfreq\_energy\_ratio} that computes $C_{0.5}$ parameter for any
input binary image is the part of \code{CorrelationFunctions.jl} package.

Our library does not have an implementation for lineal-path and chord-length
maps, although we know that such implementations exist
\cite{turner2016efficient,Havelka}. The reason is a very slow computation even
on GPUs or the resulting map is not for all points on the image, but within a
spherical volume (due to rotation of the 3D image during lineal-path
evaluation). Moreover, linear scan approach is orders of magnitude faster for
$L_2$.

Majority of described algorithms contain parallelizable operations such as fast
Fourier transform, connected components labeling and image filtering, hence they
are perfectly suited for execution on GPUs and provide very fast computations
(see execution times below).

\subsection{Directional scanning with line segment}
\label{sec:scan}
All algorithms described above (full correlation maps) are executed on the whole
input image and hence require a lot of RAM memory. The scan approach deals with
memory limitations at the cost of higher execution time (except for $L_2$). With
line-segment scanning we select a list of predefined directions and cut all
one-dimensional slices from the input along those directions. Then for each
direction, as shown in \cref{fig:directions}, we use an appropriate algorithm in
\cref{sec:map} to calculate a correlation function just for that single
slice. For this reason we do not repeat algorithms for $S_2$, $\rho_{ij}$,
$F_{ss}$ and $F_{sv}$.  When all slices are processed we calculate the resulting
directional or averaged CF using the formula:
\begin{equation*}
  F^{\bm{d}}(r) = \frac{\sum\limits_i F^{\bm{d}}_i(r) N_i(r)}{\sum\limits_i N_i(r)}
\end{equation*}
Here $F^{\bm{d}}(r)$ is a correlation function computed with respect to a
direction $\bm{d}$, $F^{\bm{d}}_i(r)$ is a ``partial'' correlation
function calculated for a $i$-th slice and $N_i(r)$ is a total number of trials
for a correlation length $r$ and slice $i$. Some algorithms require
preprocessing stages before cutting the input in slices, e.g., image filtering
for edge detection or connected components labeling performed on the whole
array, not on slices.

An algorithm for computation of linear-path function for one-dimensional slice
is presented below. Here $A$ is a one-dimensional input array, $phase$ can be
void, solid or some other phase, and $periodic$ is a boolean value which equals
to true if we compute $L_2$ in periodic mode and false otherwise. The efficiency
of the proposed algorithm is $O(n)$ where $n$ is the length of the array.
\begin{algorithmic}[1]
  \Procedure{$L_2$}{$A, phase, periodic$}
    \State $len \gets length(A)$
    \State $result \gets zeros(len)$
    \State $runs \gets countruns(A, phase)$
    \ForAll{$run \in runs$}
      \State $updateruns(result, run)$
    \EndFor
    \If{$periodic$}
      \State $first \gets first(A)$
      \State $last \gets last(A)$
      \If{$first = last$}
        \State $updateperiodic(result, first, last)$
      \EndIf
    \EndIf
    \State \textbf{return} $success$ divided by number of trials
    (see \cref{sec:number-of-trials})
  \EndProcedure
  \\
  \Procedure{countruns}{$A, phase$}
    \State $runslist \gets [\ ]$
    \State $runs \gets 0$
    \ForAll{$x \in A$}
      \If{$x = phase$}
        \State $runs \gets runs + 1$
      \ElsIf{$runs \ne 0$}
        \Comment Add $runs$ to the accumulator
        \State $runslist \gets runs:runslist$
        \State $runs \gets 0$
      \EndIf
    \EndFor
    \If{$runs \ne 0$}
      \State $runslist \gets runs:runslist$
    \EndIf
    \State \textbf{return} $runslist$
  \EndProcedure
  \\
  \Procedure{updateruns}{$A, runs$}
    \State $r \gets runs$
    \For{$idx \gets 0, runs - 1$}
      \State $A[idx] \gets A[idx] + r$
      \State $r \gets r - 1$
    \EndFor
  \EndProcedure
  \\
  \Procedure{updateperiodic}{$A, first, last$}
    \State $sum \gets first + last$
    \For{$idx \gets 0, sum - 1$}
      \State $up \gets \min(idx, first, last, sum - idx)$
      \State $A[idx] \gets A[idx] + up$
    \EndFor
  \EndProcedure
\end{algorithmic}

Cluster function can be calculated using the same technique as in map approach,
but if the number of clusters is high, a naïve $O(n^2)$ algorithm may be
preferred to $O(MC \log M)$ full map algorithm. Such an approach is present below:
\begin{algorithmic}[1]
  \Procedure{$C_2$}{$A, phase, periodic$}
    \State $A' \gets I^{(phase)}(A)$
    \State $CC \gets label\_components(A', periodic)$
    \State $len \gets length(A)$
    \State $result \gets zeros(len)$
    \ForAll{$slice \in slices(CC)$}
      \For{$i = 0,len-1$}
        \State $end \gets len-1$
        \If{$periodic$}
          \State $end \gets end + i$
        \EndIf
        \For{$j = i, end$}
          \If{$slice[i] = slice[j \mod len] \ne 0$}
            \Comment $0$ is a label for background.
            \State $dst \gets |(j \mod len) - i|$
            \State $result[dst] \gets result[dst] + 1$
          \EndIf
        \EndFor
      \EndFor
    \EndFor
    \State \textbf{return} $result$ divided by number of trials
  \EndProcedure
\end{algorithmic}

Before computation of chord length function an input array must be pre-processed
to find an interface for the phase of interest. For this purpose we use
distance transform function defined as follows:
\begin{equation}
  \mathcal{D}(\bm{x})= \left\{
  \begin{array}{ll}
    0 & \quad \bm{x} \in \text{solid phase} \\
    \min\limits_{y \in \text{solid phase}} \rho(\bm{x},\bm{y}) & \quad \text{otherwise}
  \end{array}
\right. \label{eq:distance-transform}
\end{equation}
where $\rho$ is the Euclidean distance between two points. There are many
algorithms for distance transform, including algorithms performed in linear time
\cite{DT}.In \code{chord\_length} function we firstly apply distance transform
to an input to find the interface and then accumulate chord length in per-slice
manner. Finally, chord lengths can be binned to a histogram.
\begin{algorithmic}[1]
  \Procedure{chord\_length}{$A, phase$}
    \State $A' \gets I^{(phase)}(A)$
    \State $A_E \gets \mathcal{D}(A') = 1$
    \Comment We consider elements for which the distance transform equals to $1$
    to be part of the interface
    \State $chord\_list \gets [\ ]$
    \ForAll{$(S', S_E) \in zip(slices(A'), slices(A_E))$}
      \Comment iterate through all one-dimensional slices in $A'$ and
      corresponding slices in $A_E$
      \State $l \gets 0$
      \State $maybe\_chord \gets False$
      \For{$idx \gets 0, length(S')-1$}
        \If{$S_E[idx]$}
          \Comment Interface found
          \If{$l > 1\ and\ maybe\_chord$}
            \Comment Chord found
            \State $chord\_list \gets l:chord\_list$
          \EndIf
          \State $l \gets 0$
          \State $maybe\_chord \gets True$
        \ElsIf{$\neg S'[idx]$}
          \Comment We are not in the phase of interest
          \State $maybe\_chord \gets False$
        \EndIf
        \State $l \gets l+1$
      \EndFor
    \EndFor
    \State \textbf{return} $chord\_list$ or $histogram(chord\_list)$
  \EndProcedure
\end{algorithmic}
This implementation also works in linear time.

Another function built on top of distance transform is the pore size function
$P(r)$. Recall that pore size function is a probability distribution
function. In practice, it is conveniently represented as a histogram of pore sizes
found in the input image. Our algorithm for calculating $P(r)$ is as follows:
\begin{algorithmic}[1]
  \Procedure{$P$}{$A$}
    \State $D \gets \mathcal{D}(A)$
    \Comment Apply $\mathcal{D}$ (\cref{eq:distance-transform}) to each element of $A$
    \State $D' = \{ x \in D: x \ne 0\}$
    \Comment Remove zeros
    \State Bin values in $D'$ to histogram $H$
    \State \textbf{return} $H$
    \Comment Alternatively just return $D'$
  \EndProcedure
\end{algorithmic}
Algorithmic efficiency is $O(n)$ where $n$ is the number of elements in the
input array.

\section{Verification of CFs compuations}
\label{sec:verification}
To verify all algorithms we compare their results against known closed-form
expressions for correlation functions for some particular type of input
data. One such type with analytical solution is overlapping n-balls with fixed
radius and centers generated by Poisson point process. Here we consider both
two-dimensional and three-dimensional cases.

\subsection{Two-dimensional analytical solutions}
Consider an infinite set of disks with radii $R$ and centers generated by
Poisson point process with parameter $\lambda$.

Two-point probability function in this case is \cite{Torquato_book}:
\begin{equation*}
  S_2(r) = e^{-2\lambda \left\{
  \begin{array}{ll}
    \pi R^2 & \quad r > 2R \\
    \pi R^2 + r\sqrt{A}/4 - BR^2 & \quad \text{otherwise}
  \end{array}
  \right.}
\end{equation*}
Here and later $A = 4R^2 - r^2$ and $B = \arccos \frac{r}{2R}$.

An expression for $F_{ss}$ is \cite{samarin2023robust}:
\begin{equation*}
  F_{ss}(r) = 4S_2(r) \left\{
  \begin{array}{ll}
    (\pi\lambda R)^2 & \quad r > 2R \\
    \frac{AR^2\lambda^2 r(B^2 - 2\pi B + \pi^2) + \sqrt{A}R^2\lambda}{Ar} & \quad \text{otherwise}
  \end{array}
  \right.
\end{equation*}

Surface-void function $F_{sv}$ becomes \cite{samarin2023robust}:
\begin{equation*}
  F_{sv}(r) = 2R\lambda S_2(r) \left\{
  \begin{array}{ll}
    \pi & \quad r > 2R \\
    \pi-B & \quad \text{otherwise}
  \end{array}
  \right.
\end{equation*}

Two-point function $L_2$ \cite{Torquato_book}:
\begin{equation*}
  L_2(r) = e^{-\lambda(\pi R^2 + 2rR)}
\end{equation*}

Chord length function $p$ \cite{Torquato_book}:
\begin{equation*}
  p(r) = 2\lambda R e^{-2\lambda rR}
\end{equation*}

And finally pore size function $P$ \cite{Torquato_book}:
\begin{equation*}
  P(r) = 2\lambda\pi (r+R) e^{-\lambda\pi(r^2 + 2rR)}
\end{equation*}

\subsection{Three-dimensional analytical solutions}
As in the previous case, consider an infinite set of balls with radii $R$ and
centers generated by Poisson point process with parameter $\lambda$. Closed-form
expressions for correlations functions are also known for this case for all CFs
\cite{Torquato_book}.

Two-point probability function in this case is:
\begin{equation*}
  S_2(r) = e^{-\lambda\pi \left\{
  \begin{array}{ll}
    \frac{8}{3} R^3 & \quad r > 2R \\
    \frac{4}{3} R^3 + rR^2 - \frac{1}{12}r^3 & \quad \text{otherwise}
  \end{array}
  \right.}
\end{equation*}

Let $\eta = \frac{4}{3}\lambda \pi R^3$. Then an expression for $F_{ss}$ is:
\begin{equation*}
  F_{ss}(r) = S_2(r) \left\{
  \begin{array}{ll}
    \frac{9\eta^2}{R^2} & \quad r > 2R \\
    \frac{9\eta^2}{R^2}(\frac{1}{2}+\frac{r}{4R})^2 + \frac{3\eta}{2rR} & \quad \text{otherwise}
  \end{array}
  \right.
\end{equation*}

Surface-void function $F_{sv}$ is:
\begin{equation*}
  F_{sv}(r) = \frac{3\eta}{R} S_2(r) \left\{
  \begin{array}{ll}
    1 & \quad r > 2R \\
    \frac{1}{2} + \frac{r}{4R} & \quad \text{otherwise}
  \end{array}
  \right.
\end{equation*}

Two-point function $L_2$:
\begin{equation*}
  L_2(r) = e^{-\lambda\pi (\frac{4}{3}R^3 + rR^2)}
\end{equation*}

Chord length function $p$:
\begin{equation*}
  p(r) = \pi\lambda R^2 e^{-\pi\lambda rR^2}
\end{equation*}

And finally pore size function $P$:
\begin{equation*}
  P(r) = 4\pi\lambda(r+R)^2 e^{-\frac{4}{3}\pi\lambda (r^3 + 3r^2R + 3rR^2)}
\end{equation*}

\subsection{Comparison against analytical solutions}
To test out package we generate a realization of random overlapping disks
(\cref{fig:overlapping-disks}) with dimensions $10000 \times 10000$ and
parameters of radii $R = 40\ pixels$ and $\lambda = 10^{-4}\ pixels^{-2}$ and
then compare calculated correlation functions with their theoretical values.
\begin{figure}[ht]
  \centering
  \includegraphics[width=0.45\linewidth]{images/disks.png}
  \includegraphics[width=0.45\linewidth]{images/balls.png}
  \caption[]{Images of overlapping disks and balls used for testing of our package.}
  \label{fig:overlapping-disks}
\end{figure}

For a 3D test, we generate a realization of random overlapping spheres
(\cref{fig:overlapping-disks}) with dimensions $500 \times 500 \times 500$ and
$R = 10\ voxels$ and $\lambda = 10^{-4}\ voxels^{-3}$ and then compare calculated
correlation functions with analytical solutions.

The results of comparison of computational results against analytical solutions 
are present on \cref{fig:verification}. It is clear from the test shown that CFs 
are computed with high accuracy; minor deviations are due to finite size of our
Poisson realizations (i.e., larger images reduce deviations \cite{samarin2023robust}).
The package contains automatic tests based on these comparisons.

\begin{figure*}[tp]
  \centering
  \subfigure[Lineal-path correlation function]{
    \includegraphics[width=0.3\linewidth]{images/l2.png}
    \label{fig:l2}}
  \hfill
  \subfigure[Two-point correlation function]{
    \includegraphics[width=0.3\linewidth]{images/s2.png}
    \label{fig:s2}}
  \hfill
  \subfigure[Surface-surface correlation function]{
    \includegraphics[width=0.3\linewidth]{images/ss.png}
    \label{fig:ss}}
  \vskip\baselineskip
  \subfigure[Surface-void correlation function]{
    \includegraphics[width=0.3\linewidth]{images/sv.png}
    \label{fig:sv}}
  \hfill
  \subfigure[Pore size correlation function]{
    \includegraphics[width=0.3\linewidth]{images/ps.png}
    \label{fig:ps}}
  \hfill
  \subfigure[Chord length correlation function]{
    \includegraphics[width=0.3\linewidth]{images/cl.png}
    \label{fig:cl}}
  \caption[]{A comparison of calculated and theoretical values of correlation
    functions for overlapping disks and balls.}
  \label{fig:verification}
\end{figure*}

\section{Applications and benchmarks}
After the verification of our computational approach based on analytical
solution, in this section we apply \code{CorrelationFunctions.jl} to compute CFs
for a variety of real 2D (SEM) and 3D (XCT) images. We also report benchmarks on
execution time and compare the efficiency of GPU and CPU implementations.

\subsection{Examples of application to experimental images}
\label{sec:examples}
For testing purposes we have gathered an assortment of different porous
materials including XCT images of artificial ceramics \cite{ceramics} and
natural soil composite \cite{Geoderma2018}, segmented into numerous phases, and
SEM images of oil-bearing rocks such as sandstones and carbonate
\cite{samarin2023robust}.  A summary of this library of 2D and 3D images,
including dimensions and resolutions, can be found in \cref{tab:samples}.

\begin{table}[!pt]
  \centering
  \begin{tabular}{|c|c|c|c|c|}
    \hline
    Sample name & Image type & Dimensions (pixels) & Image resolution ($\mu m$) & $C_{0.5}$\\
    \hline
    Ceramic 1   & XCT & $500 \times 500 \times 500$ & 2.25 & 0.9823 \\
    Ceramic 2   & XCT & $500 \times 500 \times 500$ & 2.25 & 0.9890 \\
    Ceramic 3   & XCT & $500 \times 500 \times 500$ & 2.25 & 0.9712 \\
    Sandstone 1 & SEM &  $1280 \times 869$ & 0.10 & 0.9703 \\
    Sandstone 2 & SEM &  $1280 \times 869$ & 0.05 & 0.9897 \\
    Carbonate 1 & SEM &  $1024 \times 691$ & 0.06 & 0.9721 \\
    \hline
    Soil (multiphase sample) & XCT & $304 \times 304 \times 304$ &
    15.0 & - \\
    \hline
  \end{tabular}
  \caption{Summary of samples used for calculation of correlation functions.}
  \label{tab:samples}
\end{table}

At first, we compare the three ceramic images which have somewhat different
genesis depending on the mixture ratio of powder and binder used for firing the
material \cite{ceramics}. As observed from \cref{fig:real-xct}, the structure of
all three samples is visually very similar. The lineal path function supports
this observation. The $S_2$ functions are also quite similar except for Sample 2
having higher porosity that results in elevated curve (the tail for two-point
probability as expected fluctuates around porosity squared). Notably, $S_2$ and
$C_2$ are almost identical -- this indicates that pore phase represents a single
large cluster. Recall that for $S_2$, $C_2$ and $L_2$ functions the value at
correlation length of 0 is the porosity value. Surface functions, on the other
hand, feel significant difference between samples in how pore-solid interface is
organized. This is further evidenced on pore-size and chord-length graphs, with
Ceramic 2 sample with the highest porosity having larger amount of large
pores. Both $P$ and $p$ show very similar behavior, but for different
correlation lengths; such behavior is attributed to quite regular pores shapes
in all ceramics samples -- this is evident from much larger differences between
pore-size and chord-length for SEM images considered next.

\begin{figure*}[pt]
  \centering
  \subfigure[Ceramic 1]{
    \includegraphics[width=0.3\linewidth]{samples/Ceramic1.png}}
  \hfill
  \subfigure[Ceramic 2]{
    \includegraphics[width=0.3\linewidth]{samples/Ceramic2.png}}
  \hfill
  \subfigure[Ceramic 3]{
    \includegraphics[width=0.3\linewidth]{samples/Ceramic3.png}}
  \vskip\baselineskip
  \includegraphics[width=0.32\linewidth]{samples/s2-xct.png}
  \includegraphics[width=0.32\linewidth]{samples/c2-xct.png}
  \includegraphics[width=0.32\linewidth]{samples/l2-xct.png}
  \vskip\baselineskip
  \includegraphics[width=0.32\linewidth]{samples/surf2-xct.png}
  \includegraphics[width=0.32\linewidth]{samples/surfvoid-xct.png}
  \vskip\baselineskip
  \includegraphics[width=0.32\linewidth]{samples/pore-size-xct.png}
  \includegraphics[width=0.32\linewidth]{samples/chord-length-xct.png}
  \caption[]{Visualization of XCT images and plots of correlation functions for
    these images.}
  \label{fig:real-xct}
\end{figure*}

The \cref{fig:real-sem} presents three 2D images of sandstones and carbonate
with computed correlation functions. In this example, unlike for quite isotropic
ceramics, we evaluated CFs in the directional manner. The $S_2$ graph
immediately provides us with porosity values. It also highlights severe
anisotropy in sandstone SEM images. For 2D images numerous patches of pores are
not connected to each other and, thus, $S_2$ and $C_2$ are very
different. Cluster functions show how connectivity decays with length, also we
can easily observe if given pore structure percolates in a given direction -
none of the images percolates (we highlighted the clusters to make this
statement easily verifiable with the naked eye). Note that $C_2$ and $L_2$ were
computed with correlations lengths equal to image dimensions explore
connectivity -- for other CFs this leads to noised curves due to smaller number
of sampled points (for example, see correlation functions' graphs in
\cite{Euras2012}).  Lineal paths are close for sandstones (with lower $L_2$
values for Sandstone 1 where large pores are occluded with clays or grain
fragments), but for carbonate with much smaller apertures of pores it is way
below $L_2$ for sandstones. Also, $L_2$ suggests that computations with
correlations lengths beyond where lineal path goes to zero are not very
informative. For very irregular pores on 2D SEM images, $P$ and $p$ functions
look much less similar compared to the ceramics case on \cref{fig:real-xct}. For
2D images it is also visually easy to analyze full CF maps, as presented on
\cref{fig:real-sem-map} for Sandstone 1 sample. Full correlation maps show
expanded version of directional functions -- asymmetry is immediately evidenced
from $S_2$, $F_{ss}$ and $F_{sv}$ maps. Moreover, the map for $C_2$ shows higher
connectivity in y-direction as compared to x-axis. The interpretation of full
correlation maps is less straightforward, but anisotropy patterns in directions
different from those presented in \cref{fig:directions} can be spotted.

\begin{figure*}[pt]
  \centering
  \subfigure[Sandstone 1]{
    \includegraphics[width=0.3\linewidth, frame]{samples/sandstone1.png}}
  \hfill
  \subfigure[Sandstone 2]{
    \includegraphics[width=0.3\linewidth, frame]{samples/sandstone2.png}}
  \hfill
  \subfigure[Carbonate 1]{
    \includegraphics[width=0.3\linewidth, frame]{samples/carbonate1.png}}
  \vskip\baselineskip
  \includegraphics[width=0.32\linewidth]{samples/s2-sem.png}
  \includegraphics[width=0.32\linewidth]{samples/c2-sem.png}
  \includegraphics[width=0.32\linewidth]{samples/l2-sem.png}
  \vskip\baselineskip
  \includegraphics[width=0.32\linewidth]{samples/surf2-sem.png}
  \includegraphics[width=0.32\linewidth]{samples/surfvoid-sem.png}
  \vskip\baselineskip
  \includegraphics[width=0.32\linewidth]{samples/pore-size-sem.png}
  \includegraphics[width=0.32\linewidth]{samples/chord-length-sem.png}
  \caption[]{SEM images and plots of correlation functions for these
    images. Different clusters of the void phase are shown in color.}
  \label{fig:real-sem}
\end{figure*}

\begin{figure*}[pt]
  \centering
  \subfigure[$S_2$]{
    \includegraphics[width=0.45\linewidth]{samples/map-s2.png}}
  \hfill
  \subfigure[$C_2$]{
    \includegraphics[width=0.45\linewidth]{samples/map-c2.png}}
  \vskip\baselineskip
  \subfigure[$F_{ss}$]{
    \includegraphics[width=0.45\linewidth]{samples/map-surf2.png}}
  \hfill
  \subfigure[$F_{sv}$]{
    \includegraphics[width=0.45\linewidth]{samples/map-surfvoid.png}}
  \caption[]{Correlation maps for Sandstone 1.}
  \label{fig:real-sem-map}
\end{figure*}

Finally, we conclude this section with the example of multi-phase soil composite
for which we computed cross-correlation functions \cref{fig:real-multiphase}.
The graph on the right side of the figure shows cross-correlations between
different pairs of phases (normalized by each phase's ratio). One can easily
conclude that organo-mineral complex (phase 2) is mainly located on pore (phase
1) walls. Heavy minerals (phase 4) do not cross-correlate with pores, as they
are mainly lie within the mineral solid phase (phase 3) -- they do show high
cross-correlations. Other CFs can be easily computed for each separate phase,
while $F_{sv}$ can potentially be extended beyond void (or pores) and be
represented by any other phase or their superposition.

\begin{figure}[ht]
  \centering
  \includegraphics[width=0.45\linewidth]{samples/multiphase.png}
  \includegraphics[width=0.45\linewidth]{samples/multiphase-cc.png}
  \caption[]{Cross-correlation functions for multiphase sample (soil) which has
    4 distinct phases, labeled as 1-4. Each function $\rho_{ij}$ is normalized
    by division by $\rho_{ii}(0) \rho_{jj}(0)$.}
  \label{fig:real-multiphase}
\end{figure}

This section clearly demonstrated, that \code{CorrelationFunctions.jl} can be
utilized in analysis of any 2D and 3D images, including structures consisting of
numerous phases.

\subsection{Execution times for our algorithms}
\label{sec:efficiency}
To measure execution times for developed computational algorithms we use a
machine with the following configuration:
\begin{itemize}
\item \textbf{CPU}: AMD Ryzen 7 5800X
\item \textbf{Memory}: 32~GB DDR4 memory
\item \textbf{GPU}: Nvidia Tesla V100 with 32~GB video memory
\end{itemize}

We generate 2 and 3-dimensional square/cubical binary arrays with different
width (see insets on \cref{fig:timings}). An array used for measurements
contains two solid disks or balls surrounded by void phase. This configuration
is a compromise between ease of image scaling and complexity of the structure
(e.g., full map $C_2$ computational time is not compromised by too many
clusters).

We generated six arrays with dimensions of $s = 1000, 2000, \dots, 6000$ pixels
for two-dimensional case and six arrays with a side of a cube $s = 50, 100,
\dots, 300$ voxels for three-dimensional case and calculated execution time for
each correlation function. All functions are calculated using periodic boundary
conditions. Functions which use per-slice directional approach were calculated
along major axial directions only.

Functions which calculate full correlation maps are suitable for running on both
CPU and GPU, so we provide execution times for both. The cluster function uses
an algorithm for closest component labeling which has CPU-only implementation
(thus, leaving a room for future improvements). Directional computations are
performed exclusively on CPUs.

All measurements of execution time are presented on \cref{fig:timings}. It is
immediately apparent that GPU versions (where available) are much faster than
CPU versions and limited only by amount of video memory. Functions in module
\code{Directional} (which utilize per-slice directional approach) are the
slowest among other implementations, but require less memory.

\begin{figure*}[t]
  \centering
  \subfigure[Two-dimensional case, module \code{Directional}]{
    \includegraphics[width=0.475\linewidth]{images/time-2d.png}
    \label{fig:timings-2d}}
  \hfill
  \subfigure[Three-dimensional case, module \code{Directional}]{
    \includegraphics[width=0.475\linewidth]{images/time-3d.png}
    \label{fig:timings-3d}}
  \vskip\baselineskip
  \subfigure[Two-dimensional case, module \code{Map}, CPU]{
    \includegraphics[width=0.475\linewidth]{images/time-2d-map.png}
    \label{fig:timings-2d-map}}
  \hfill
  \subfigure[Three-dimensional case, module \code{Map}, CPU]{
    \includegraphics[width=0.475\linewidth]{images/time-3d-map.png}
    \label{fig:timings-3d-map}}
  \vskip\baselineskip
  \subfigure[Two-dimensional case, module \code{Map}, GPU]{
    \includegraphics[width=0.475\linewidth]{images/time-2d-gpu.png}
    \label{fig:timings-2d-gpu}}
  \hfill
  \subfigure[Three-dimensional case, module \code{Map}, GPU]{
    \includegraphics[width=0.475\linewidth]{images/time-3d-gpu.png}
    \label{fig:timings-3d-gpu}}
  \caption[]{Execution times for different correlation functions.}
  \label{fig:timings}
\end{figure*}

\section{Discussion and outlook}
\label{sec:outlook}
In the light of the end of the previous section it is important to clarify that
available (video) RAM or speed of computations are not the only considerations
behind choosing between full correlation map and directional linear
scans. Directional CFs are extremely useful in describing or reconstructing
anisotropic structures \cite{jiao2014chawla,EPL1,tensorPRE}. At the same time,
they also much easier to work with and good candidates for data compression and
feature extraction \cite{KarsaninaEJSS}. On the other hand, full map for $S_2$
allows computing ensemble averaged two-point probability -- in the similar
fashion it is measured with scattering experiments. The usefulness of full
correlation maps for other CFs is questionable, as they are either
non-measurable experimentally, e.g., \cite{Gommes_notchord}, or non-applicable
for stochastic reconstructions.  The latter refers to our ability to reconstruct
almost any structure from its $S_2$ map \cite{fullwood2008}. And while we have
fast and efficient techniques such as phase-recovery to reconstruct the
structure from its $S_2$ correlation map, the utilization of other CFs is
complicated \cite{cherkasov2021,cherkasovIC}. Simulated-annealing based methods
that operate on the full map are, to the best of our knowledge, were not
developed to date. This is the reason we did not invest into improvement on
existing approaches to compute maps for $L_2$.

The trade-off in using full map or directional CFs is closely related to the
information content of correlation functions
\cite{Gommes1,Gommes2,skolnick2021}. Any reduction of the full map into ensemble
averaged or directionally computed CFs significantly decreases the information
available. On the other hand, full correlation map is larger than the image
itself and, thus, storing such huge amount of data for already voluminous
datasets is not practical.  Vectorization of directional correlation functions
is, thus, seems like a natural feature reduction approach while working with
correlation maps. As applied to stochastic reconstructions, similar logic
applies - it seems to be beneficial to increase the number of functions rather
than their "volume". Full correlation maps are usually not available unless the
structure is already known; but they may be useful in the form of 2D maps
available from images for 3D reconstructions \cite{cherkasov2021}. Wether
reconstructions of, say, full 3D maps based on 2D maps is a better approach than
using directional CFs computed from images for stochastic reconstructions is an
open question. Simulated annealing based methods were developed in the first
place due to absence of methods like phase-recovery for correlation functions
other than autocorrelation $S_2$. We believe that our fast and robust
computational framework in the form of \code{CorrelationFunctions.jl} package
will facilitate research in these directions.

While the toolbox of presented package and its API are solid, based on
additional research numerous areas for future improvements of
\code{CorrelationFunctions.jl} are possible:
\begin{enumerate}
\item include novel non-classical correlation functions, e.g., describing
  topology of phases;
\item implementing sub-cube based computations of CFs with a given correlation
  functions with later averaging for larger volume - with such domain
  decomposition one could compute CFs for larger samples with limited GPU
  resources;
\item implementing GPU-based clustering (cluster labeling) and distance-map
  transform to further speed up computations for $C_2$, chord-length and
  pore-size functions;
\item implementing higher order functions, such as 3-point probabilities or even
  3-point cluster and other more involved CFs;
\item increase number of directions for scan approach or implementing
  calculations within a given solid angle.
\end{enumerate}
We perform steps forward in some directions, whilst others still require
additional research to be made to facilitate their implementation within the
package.

\section{Summary}
\label{sec:summary}
In this paper we introduced \code{CorrelationFunctions.jl} open-source package
developed in Julia and capable of computing all classical correlation functions
based on 2D and 3D input images. Our code is capable of evaluating two-point
probability, cross-correlation, cluster, lineal-path, surface-surface,
surface-void, pore-size and chord-length functions on both CPU and GPU
architectures. There possible, we presented two types of computations: full
correlation map (correlations of each point with other points on the image, that
also allows obtaining ensemble averaged CF) and directional correlation
functions (currently in major orthogonal and diagonal directions). Such an
implementation allowed for the first time to assemble a completely free solution
to evaluate correlation functions under any operating system with well
documented API. Our package includes automatic tests against analytical
solutions that are described in this paper. We measured execution times for all
CPU and GPU implementations and as a rule of thumb full correlation maps are
faster than directional scans because calculation of correlation maps applies
fast Fourier transform to the whole input array while, when calculating scans,
FFT needs to be applied to each scan separately. Calculation on GPU shows
performance increase when compared to CPU, as expected.  However, full maps
require more RAM and, thus, are limited to available RAM resources. On the other
hand, directional CFs are memory efficient and can be evaluated for huge
datasets -- this way they are the first candidates for structural data
compression of feature extraction. The package itself is available through Julia
package ecosystem and on GitHub
(\url{https://github.com/fatimp/CorrelationFunctions.jl}), the latter source
also contains documentation and additional helpful resources such as tutorials.
We believe that a single powerful computational tool such as
\code{CorrelationFunctions.jl} will significantly facilitate the usage of
correlation functions in numerous areas of structural description and research,
as well as in machine learning applications.

\section{Acknowledgments}
This research was supported by the Russian Science Foundation grant 19-72-10082.

Collaborative effort of the authors within the FaT iMP (Flow and Transport in
Media with Pores) research group (www.porenetwork.com) and used some of its
software.

\appendix
\section{Computation of number of trials for non-periodic mode}
\label{sec:number-of-trials}
Here we assume that an input is padded with zeros to the length $2s_i - 1$
across each dimension $i$ where $s_i$ is the size of the non-padded input in
that respective dimension. For each dimension we calculate a sequence:
\begin{equation*}
  N^i_j = \left\{
  \begin{array}{cc}
    s_i - j & \quad j \in \overline{0, s_i-1} \\
    j - s_i + 1 & \quad j \in \overline{s_i, 2s_i-2}
  \end{array}
  \right.
\end{equation*}

When a number of trials is calculated as follows:
\begin{equation*}
  N_{i_1, i_2, \dots, i_n} = \prod_{k=1}^n N^k_{i_k}
\end{equation*}
Here $n$ is a number of dimensions and $i_k$ is an index into the output array
for $k$-th dimension.

\bibliography{apssamp}
\end{document}